---
title: "MLperformance"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ML_Performance}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction

MLperformance is a package that allows the user to calculate the performance metrics of a machine learning classifier from a dataset of two binary variables: actual and predicted. It creates a confusion matrix and calculates prevalence dependent and prevalence independent metrics. The dataset called "barretts" is included for an example.

cmcounts() is a function that calculates the inner cells of the confusion matrix.  Save as a cmcounts object.

conmat() creates the confusion matrix from the cmcounts object or from a vector of frequencies of binary outcomes: 00,01,10,11. Save as a confusion matrix object.

metrics() calculates the confusion matrix cells and margins and outputs in a dataframe. (P)-Total positive, (N)-Total negative, (PP)-Predicted positive, (PN)-Predicted negative, (TP)-True positive, (TN)-True negative, (FP)- False positive, (FN)-False negative, (n)-sample size

pdm() calculates prevalence dependent metrics. Prevalence, (PPV)- positive predictive value, (NPV)- negative predictive value, (FDR)- false discovery rate, (FOR)- false omission rate, (MCC), Matthews correlation coefficient

pim() calculates prevalence independent metrics. (ACC)-Accuracy, (TPR)- True Positive rate, "Sensitivity", (TNR)- True Negative Rate, "Specificity", (BA)- Balanced Accuracy, (BM)- Bookmaker informedness, (FPR)- False Positive Rate, (FNR)- False Negative Rate, (PLR)- Positive Likelihood Ratio, (NLR)- Negative Likelihood Ratio, (DOR)- Diagnostic Odds Ratio

conf() calculates the 95% confidence interval of the diagnostic odds ratio.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Function examples with barretts data

```{r setup}
library(MLperformance)
```

The dataset called "barretts" included with this R package, contains two binary variables: actual and predicted.  A machine learning classifier was trained on H&E pathology slides on patients diagnosed with Barrett's esophagus containing no dysplasia.  There was a 10-year follow-up of patients to identify development of any level of dysplasia (known as "progressors") versus patients who did not develop any dysplasia (known as "non-progressors"). The variable values denote non-progressors=0 and progressors=1.  The machine learning classifier used 90% of cases as training cases to predict progression in the remaining 10% used as test cases. This dataset will be used with the MLperformance R package to calculate confusion matrix metrics to assess the performance of the machine learning classifier.

The dataset "barretts" is loaded into the global environment and the cmcounts() function is used to calculate the cells of the confusion matrix. The cmcounts results are saved as a counts object.

```{r}
library(MLperformance)
counts <- cmcounts(barretts)
```

The conmat() function is used with the cmcounts object to create the confusion matrix, and saved as a confusion matrix object.

```{r}
cm <- conmat(counts)
cm
```


Alternatively, a vector of binary outcomes can be used to create the confusion matrix: 

```{r}
cm <- conmat(c(43,20,5,31)) #frequencies of binary outcomes: 00,01,10,11
```

The confusion matrix object can now be used to calculate various metrics. Using the metrics() function, the confusion table metrics and margins are calculated and can be saved in a dataframe.

```{r}
metrics(cm)
```


The pim() function is used to calculate the prevalence independent metrics from the confusion matrix object. This can also be saved in a dataframe.

```{r}
pim(cm)
```

The pdm() function is used to calculate the prevalence dependent metrics from the confusion matrix object.  

```{r}
pdm(cm)
```

The 95% confidence interval of the diagnostic odds ratio can be calculated using the conf() function.

```{r}
conf(cm)
```